---
output: github_document
---

<!-- TODO README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ModelArray

<!-- badges: start -->
[![CircleCI build status](https://circleci.com/gh/PennLINC/ModelArray.svg?style=svg)](https://circleci.com/gh/PennLINC/ModelArray)
<!-- badges: end -->

`ModelArray` is a generalizable, memory-efficient R package for statistical analysis of fixel data. Its features include:

* Easy to use: set up your analysis with just several lines of codes;
* Low memory requirement, even for large datasets;
* At present, `ModelArray` supports linear models as well as generalized additive models (GAM) with penalized splines, which are particularly useful for studying nonlinear effects in lifespan data.  


## Installation
Please refer to webpage [Installation](https://pennlinc.github.io/ModelArray/articles/a01_installations.html) for a full guidance of installation of `ModelArray` and its companion python package [ConFixel](https://github.com/PennLINC/ConFixel). The most important steps for installing `ModelArray` are:

* Make sure you have necessary libraries for HDF5 - see [this section](https://pennlinc.github.io/ModelArray/articles/a01_installations.html#install-hdf5-libraries-in-the-system)
* Install `ModelArray` from GitHub - see [this section](https://pennlinc.github.io/ModelArray/articles/a01_installations.html#install-modelarray-r-package-from-github)

## How to use
Load the `ModelArray` package into R via:
```{r example}
library(ModelArray)
```

We provide a example walkthrough [here](https://pennlinc.github.io/ModelArray/articles/a02_walkthrough.html). 

For documentations of `ModelArray` functions, you can:

* Either go to [this webpage](https://pennlinc.github.io/ModelArray/reference/index.html); 
* Or in R console, type: `help(<function_name>)`. For example: `help(ModelArray.lm)` 



```{r, include = FALSE}
# 1. Confirm links above are up-to-date!!
# 2. You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

# You can also embed plots, for example:

# ```{r pressure, echo = FALSE}
# plot(pressure)
# ```

# In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.


```




