---
output: github_document
---

<!-- TODO README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ModelArray

<!-- badges: start -->
[![CircleCI build status](https://circleci.com/gh/PennLINC/ModelArray.svg?style=svg)](https://circleci.com/gh/PennLINC/ModelArray)
<!-- badges: end -->

ModelArray is a generalizable, memory-efficient R package for statistical analysis of fixel data. Its features include:

* Easy to use: set up your analysis with just several lines of codes;
* Low memory requirement, even for large datasets;
* At present, ModelArray supports linear models as well as generalized additive models (GAM) with penalized splines, which are particularly useful for studying nonlinear effects in lifespan data.  


## Installation
Before you install ModelArray R package, if you are using Linux system, please check if libhdf5-dev has been installed in your system:
```console
foo@bar:~$ ldconfig -p | grep libhdf5*
```
If you got more than one line of outputs, congrats, you have libhdf5-dev installed. Otherwise, please install it. For Ubuntu user, you may install via:
```console
foo@bar:~$ sudo apt-get update -y
foo@bar:~$ sudo apt-get install -y libhdf5-dev
```

You can install the latest version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("PennLINC/ModelArray")
```
## How to use
Load the package into R via:
```{r example}
library(ModelArray)
```
We provide a quick start guide [here](https://pennlinc.github.io/ModelArray/articles/ModelArray_basics.html). \
For documentations of functions, you can either refer to [this webpage](https://pennlinc.github.io/ModelArray/reference/index.html); 
Or in R Studio console: `?<function_name>`. For example: `?ModelArray.lm()` \
For users who are not familiar with R, you may check out [this webpage](https://pennlinc.github.io/ModelArray/articles/basic_r_intro.html). 



```{r, include = FALSE}
# 1. Confirm links above are up-to-date!!
# 2. You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

# You can also embed plots, for example:

# ```{r pressure, echo = FALSE}
# plot(pressure)
# ```

# In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.


```




