---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This documentation walk through the installation of `ModelArray` package and necessary dependent packages/libraries.

Currently there are two ways to install and use `ModelArray`: 

* Directly installing from [GitHub](https://github.com/PennLINC/ModelArray)
  * This is only fully tested and supported on Linux Ubuntu system and macOS system. If you are a Windows user, you will face errors when running with more than one CPU cores (`n_cores = 2 or above`). You may install a Linux subsystem or virtual machine on your Windows machine.
* Using ModelArray container
  * Just like usual packages, ModelArray also depends on other packages/libraries including the companion python package `ConFixel`. A ModelArray container has wrapped all these dependencies and it is ready to use. This is suitable for running ModelArray on a cluster.



## Directly installing from GitHub
We will first set up the environment, then install some dependent packages and libraries, and finally install ModelArray.

### Set up a conda environment

We first create a conda environment `modelarray` for installing `ConFixel` etc. We'll install python version 3.9:

``` {.console}
foo@bar:~$ conda create --name modelarray python=3.9
foo@bar:~$ conda activate modelarray
```

### Install MRtrix

`ConFixel` uses function `mrconvert` from MRtrix, so please make sure MRtrix has been installed. It can either be installed via `conda` in this conda environment we just created, or be compiled from source. See [MRtrix's webpage](https://www.mrtrix.org/download/) for more. Type `mrview` in the terminal to check whether MRtrix installation is successful.

### Install hdf5 libraries in the system

Because ModelArray works with HDF5 file format, we need to make sure necessary libraries of HDF5 are installed in the system.

#### On a Linux Ubuntu system

If you're on Linux Ubuntu system: First, please check if libhdf5-dev has been installed in the system:

``` {.console}
foo@bar:~$ ldconfig -p | grep libhdf5*
```

If you got more than one line of outputs, congrats, you have libhdf5-dev installed. Otherwise, please install it via:

``` {.console}
foo@bar:~$ sudo apt-get update -y
foo@bar:~$ sudo apt-get install -y libhdf5-dev
```

#### On a macOS system

Use `Homebrew` to install:

``` {.console}
foo@bar:~$ brew install hdf5
```

For details you may refer to the webpage [here](https://formulae.brew.sh/formula/hdf5)

### Install ConFixel python package from GitHub
``` {.console}
# We first activate the conda environment we just created:
foo@bar:~$ conda activate modelarray  

# Then install ConFixel:
foo@bar:~$ cd ~/myProject
foo@bar:myProject$ git clone https://github.com/PennLINC/ConFixel.git
foo@bar:myProject$ cd ConFixel
foo@bar:myProject$ pip install .

# You may remove the original source code if you are an end user instead of not a developer:
foo@bar:myProject$ cd ..
foo@bar:myProject$ rm -r ConFixel
```


### Install R (>=4.1.2)

`ModelArray` requires R >=4.1.2. If you have already installed it, you can skip this step. If you don't, you may download it from [CRAN](https://cran.r-project.org/).

### (Optional) Install RStudio

RStudio provides a good IDE for using R. However it's optional to install RStudio. If you haven't got one but want to install it, you may download it from [here](https://www.rstudio.com/products/rstudio/download/).

### Install ModelArray R package from GitHub

ModelArray R package's source code is available on GitHub. To install it in R:
```{r install ModelArray, eval=FALSE}
# First, load library "devtools":
library(devtools)

# if you got error, it means you don't have devtools installed, please install it by: install.packages("devtools")

# Then, install ModelArray:
devtools::install_github("PennLINC/ModelArray")
```

Now, ModelArray is ready to use:
```{r}
library(ModelArray)
```


## Using ModelArray container
This container not only includes ModelArray, but also ConFixel and dependent packages/libraries.

You can pull the docker image from xxxxxx
TODO............
