---
title: "An intro to R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An intro to R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(ModelArray)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

***UNDER DEVELOPMENT...***

This document is a quick introduction to basic R. Audience would be users who are not familiar with R.

## R syntax

* Dot "." is valid in a variable name or a function (e.g. variable "num.fixels", function "ModelArray.lm()")
* Syntax for formula: 
Below is an example formula:
```{r}
formula <- FD ~ age + sex
```
** As you can see we did not define variable "FD","age" or "sex", but it's a valid formula. 
** Left hand side of "~" is dependent variable (y, or response), and righ hand side includes the independent variables and covariates


## Prepare the phenotypes data.frame: manipulations
Users may want to manipulate the data.frame `phenotypes` before passing it to `ModelArray.lm()` etc functions for model fitting. Examples like de-mean and rescale covariates. R package "dplyr" is useful and easy to use for manipulation. 
```{r}
# install is first via: install.packages("dplyr")
library(dplyr)    # load into memory
```

<!---
### De-mean and rescale covariates
TODO

### Check whether the model is well-conditioned
TODO

--->
