---
title: "Developer documentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developer documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(ModelArray)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Target audience: developers, and anyone from the broader community who hopes to contribute to `ModelArray`.

In this documentation, we walk through important methods used in `ModelArray` package, and the structure of the package and its functions.

We welcome community contributions to `ModelArray`. If you hope to improve functions or fix a bug in our package, please follow the following steps:

* fork the [GitHub repository](https://github.com/PennLINC/ModelArray)
* modify the scripts
* update the tests and/or add necessary new tests (see folder `tests`)
* run tests + update docs and website:
  * to make sure the unit tests and package building are without error;
  * to make sure the documentations and website can be built, and to update any documentations;
  * to do so, run the following commands: `devtools::build_readme()`, `devtools::check()`, `devtools::build()`, `devtools::build_vignettes()`, and finally `pkgdown::build_site()`. Please check whether there is no error along the process, and the built website looks good.
* if everything looks good, please submit a pull request to our github repository.

## Overview

`ModelArray` R package supports data in Hierarchical Data Format 5 (HDF5) file format. It utilizes DelayedArray and HDF5Array R packages, so that statistical analysis can be performed without loading all the original data in HDF5 file into memory. This makes `ModelArray` very efficient in memory usage. For details, please see section "ModelArray Construction" below. \

Methods for showing and accessing `ModelArray` object are basic. Please see section "Show and accessors methods of ModelArray class" for more.

The key function of `ModelArray` package is to perform statistical analysis. For details, please see section "Model fitting" below. \

Then we describe how to save statistical results to HDF5 file on disk.

<!-- deal with missing value, -->

Furthermore, we describe what unit tests are needed for quality assurance.

Finally we describe building the Docker image.

## `ModelArray` Construction
We first define a class called `ModelArray` using `setClass()`. An `ModelArray` object (i.e. an instance of `ModelArray` class) has several slots:

* sources: source filenames (e.g., a list of .mif files of fixel-wise data)
* scalars: scalar matrix (or matrices)
* results: statistical result matrices (if any)
* path: the path to the h5 file on disk

The key feature of an `ModelArray` object is memory efficient. This is because the entire dataset in HDF5 (.h5) file was not loaded into the memory; only minimal data was loaded. To achieve this, we first need `ModelArraySeed()`, which utilizes `HDF5Array::HDF5ArraySeed()`, acting as a pointer to the .hdf5 file on disk. To make the arrays in the `ModelArray` object look more like "real" arrays, e.g. common array operations such as indexing and transposing can be applied, we utilize package `DelayedArray` to wraps the data in on-disk HDF5 file into a `DelayedArray` object. Finally a `ModelArray` class is defined by integrating above slots together. This is done in the `ModelArray()` function.

Above functions and setups can be found in script `R/ModelArray_Constructor.R`

## Show and accessors methods of `ModelArray` class
To easily show and access a `ModelArray` object, we adopt S4 Object Oriented Programming (OOP) model.

End users of R actually are using such S4 methods frequently. For example, when showing the (summarized) content of an object, users simply enter its name in R console and press "enter". Under the hood, it uses the generic `show()` method. As for showing a `ModelArray` object, we adopt the same way. As `show()` has already been a generic function, we only need to `setMethod()` for its setup.

With S4 OOP model, users can easily access the data they need. Below are current accessor functions for accessing different slots in `ModelArray` class's:

* `sources()`: source filenames (e.g., a list of .mif files of fixel-wise data)
* `scalars()`: scalar matrix (or matrices)
* `results()`: statistical result matrix or matrices (if any)

We use `setMethod()` to define above methods. Unlike `show()`, these accessors are not generic yet, so we then do additional step of `setGeneric()`.

All of these show and accessors setups can be found in script `R/ModelArray_S4Methods.R`.


## Model fitting
Functions such as `ModelArray.lm()` and `ModelArray.gam()` are for model fitting. Under the hood, these functions iteratively call `analyseOneElement.lm()` and `analyseOneElement.gam()` for fitting for one element, respectively. This also facilitates the parallel computing across all elements requested.

Besides the existing functions for linear models (`ModelArray.lm()`) and GAMs (`ModelArray.gam()`), new functions for diverse statistical models can be added by following the structure below.

The general structure for model fitting functions `ModelArray.<model_name>()` in `ModelArray` is as below:

* Some sanity checks of arguments
* Print out some important methods for model fitting, such as `method` in `mgcv::gam()` for GAMs
* Fit statistical models:
  * First, initialize: run `analyseOneElement.<model_name>()` once to get the column names of the statistics in the output statistics; usually using an element in the middle (instead of the first or the last element) to avoid too many NaNs in the data;
  * Then, iteratively call `analyseOneElement.<model_name>()` across all elements requested (can be parallelized). The function `analyseOneElement.<model_name>()` does the following jobs:
    * Fit the model for this element;
    * Get necessary statistics via `broom::tidy()`, `broom::glance()`, and/or `summary()`.
    * Flatten the results into one row of data.frame for this element; add element ids; remove the column names (it's a list of numeric now) to save a bit more memory
    * Return this result of this element
  * Concatenate results of all requested elements into a matrix of elements by statistics, then add the column names; now we get the final result data.frame
* If _p_-value correction (for terms and/or model) is requested:
  * For each _p_-value correction method, apply it and add the values to the result data.frame
* Return the result data.frame

For `ModelArray.gam()`, there is an additional step after one round of iteration:

* If `changed.rsq.term.index` is not `NULL`, i.e. there is request of quantifying one (or more terms)'s importance in the model (i.e. metrics delta adjusted R-squared and partial R-squared), we need to compute the reduced model (formula without the term of interest):
  * For each requested term of interest:
    * Iteratively call `analyseOneElement.<model_name>()` but providing reduced formula
    * Get necessary statistics
    * Compute delta adjusted R-squared and partial R-squared; add them to the previously got data.frame

Therefore, there might be two or more rounds of iterations on model fitting in `ModelArray.gam()`, depending on number of terms the user requested in `changed.rsq.term.index`.

Functions mentioned in this section can be found in scripts `R/analyse.R` and `R/ModelArray_Constructor.R`.


## Saving statistical results to .hdf5 file on disk
After getting the data.frame of statistical results, users will save it into the original HDF5 file on disk. The function is `writeResults()` located in script `R/ModelArray_Constructor.R`

<!--
## Dealing with missing values
### Missing values in `phenotypes` data.frame
TODO

### Missing values in `ModelArray` scalar data
TODO
-->

## Writing tests for CircleCI
We use Continuous Integration (CI) testing to ensure stability and quality assurance of `ModelArray`. Therefore, when you modify the source code of `ModelArray`, it is very important to make sure you add appropriate tests of the new or modified functionality.

All the test files are located in folder `tests/testthat`. Once a commit is pushed to GitHub, CircleCI (a CI testing platform) will be triggered to build the software and run the unit tests.

The tests serve the goals of:

1. Run functions in different scenarios (with different requests) and expect no error. This is to ensure when new features added, it will run without error;
  ** For some scenarios, some special data is needed. For example, when testing NAs in input data (for volume data)
2. Expect the output data.frame will contain column with a specific name, or check the dimensions of the output data.frame. This is to ensure expected columns show up.
3. Most importantly, check the values are as expected: statistical results calculated in ModelArray fitting loop = those calculated in standard R.

Developers can run the tests locally to see if unit tests in a test file (in `tests/testthat` folder) can run without errors:

```{r, eval=FALSE}
rm(list=ls())
devtools::load_all()
```

Then click "Run Test" button in RStudio to run the test file. Check if there is anything failed.

For more details on how to write unit tests + test out, please check out `Testing` chapters in the book ["R Packages" written by Hadley Wickham, Jennifer Bryan](https://r-pkgs.org/index.html)

## Building the Docker image for ModelArray + ConFixel
The `Dockerfile` for the Docker image can be found in the root folder of the GitHub repository of `ModelArray`.

CircleCI will automatically build the Docker image and push to Docker Hub when there is a GitHub commit or a merge to main branch, or when a version is tagged. See the `.circleci/config.yml` file for more details on these setups.


