---
title: "Splitting Element IDs for Parallel Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Splitting Element IDs for Parallel Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Why split elements across jobs?

ModelArray supports within-process parallelism via the `n_cores` parameter -
multiple CPU cores in a single R session working through elements in parallel.
For many datasets, this is sufficient.

However, for very large datasets (e.g., 600,000+ fixels with thousands of subjects), 
even a multi-core single job may take too long. 
In these cases, you can **split the elements across multiple independent HPC jobs**, 
each processing a different range of elements.

## Generating split ranges

Use `numElementsTotal()` to get the total element count, then divide it into ranges:

```{r split_ranges, eval=FALSE}
library(ModelArray)

modelarray <- ModelArray("data.h5", scalar_types = c("FDC"))
n_total <- numElementsTotal(modelarray, "FDC")

# Split into chunks of 100,000 elements
chunk_size <- 100000
starts <- seq(1, n_total, by = chunk_size)
ends <- pmin(starts + chunk_size - 1, n_total)

# Each job gets one of these ranges
for (i in seq_along(starts)) {
  cat(sprintf("Job %d: element.subset = %d:%d\n", i, starts[i], ends[i]))
}
```

```{.console}
Job 1: element.subset = 1:100000
Job 2: element.subset = 100001:200000
Job 3: element.subset = 200001:300000
Job 4: element.subset = 300001:400000
Job 5: element.subset = 400001:500000
Job 6: element.subset = 500001:600000
Job 7: element.subset = 600001:602229
```

## Running each job

Each job runs the same model on its assigned range of elements. 
For example, an R script parameterized by job index:

```{r job_script, eval=FALSE}
# run_chunk.R
# Usage: Rscript run_chunk.R <job_index>

args <- commandArgs(trailingOnly = TRUE)
job_index <- as.integer(args[1])

library(ModelArray)

modelarray <- ModelArray("data.h5", scalar_types = c("FDC"))
phenotypes <- read.csv("cohort.csv")

chunk_size <- 100000
n_total <- numElementsTotal(modelarray, "FDC")
start <- (job_index - 1) * chunk_size + 1
end <- min(job_index * chunk_size, n_total)

result <- ModelArray.lm(
  FDC ~ Age + sex + motion,
  modelarray, phenotypes, "FDC",
  element.subset = start:end,
  n_cores = 4
)

# Save this chunk to its own HDF5 file
# (safe because each job writes to a different file)
chunk_h5 <- sprintf("result_chunk_%d.h5", job_index)
writeResults(chunk_h5, df.output = result, analysis_name = "results_lm_chunk")
```

On an HPC cluster with a job array (e.g., SLURM):

```{.console}
#!/bin/bash
#SBATCH --array=1-7
#SBATCH --cpus-per-task=4

Rscript run_chunk.R $SLURM_ARRAY_TASK_ID
```

## Combining results and writing to HDF5

After all jobs complete, read each chunk `.h5`, concatenate rows, then write one final result to `data.h5`:

```{r combine_results, eval=FALSE}
library(ModelArray)
library(rhdf5)

# Collect chunk HDF5 files and order by numeric chunk index
chunk_files <- Sys.glob("result_chunk_*.h5")
if (length(chunk_files) == 0) {
  stop("No chunk HDF5 files found (expected files like result_chunk_1.h5).")
}
chunk_ids <- as.integer(sub("result_chunk_([0-9]+)\\.h5$", "\\1", basename(chunk_files)))
if (anyNA(chunk_ids)) {
  stop("Could not parse chunk index from one or more chunk filenames.")
}
chunk_files <- chunk_files[order(chunk_ids)]

# Read each chunk's saved results matrix + column names
results_list <- lapply(chunk_files, function(fn) {
  mat <- h5read(fn, "results/results_lm_chunk/results_matrix")
  col_names <- h5read(fn, "results/results_lm_chunk/column_names")
  df <- as.data.frame(mat)
  colnames(df) <- as.character(col_names)
  df
})

# Optional consistency check across chunks
if (!all(vapply(results_list[-1], function(x) identical(colnames(x), colnames(results_list[[1]])), logical(1)))) {
  stop("Column names differ across chunk files; cannot concatenate safely.")
}

full_result <- do.call(rbind, results_list)

# Verify we have all elements
cat("Total rows:", nrow(full_result), "\n")

# Single-process write back to the main HDF5 file
writeResults("data.h5", df.output = full_result, analysis_name = "results_lm")
```

**Important**: only one process should write to the HDF5 file at a time. HDF5 does not support concurrent writes from multiple processes — see `vignette("hdf5-large-analyses")` for details on why.
