---
title: "Exploring HDF5 Data with Convenience Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring HDF5 Data with Convenience Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

ModelArray provides several accessor functions for inspecting your data without writing analysis code.
This vignette walks through each one.

```{r setup, eval=FALSE}
library(ModelArray)

h5_path <- "~/Desktop/myProject/demo_FDC_n100.h5"
modelarray <- ModelArray(h5_path,
  scalar_types = c("FDC"),
  analysis_names = c("results_lm")
)
phenotypes <- read.csv("~/Desktop/myProject/cohort_FDC_n100.csv")
```

## `show()` — Object summary

Simply typing the object name (or calling `show()`) prints a summary:

```{r show, eval=FALSE}
modelarray
```

```{.console}
ModelArray located at ~/Desktop/myProject/demo_FDC_n100.h5

  Source files:     100
  Scalars:          FDC
  Analyses:         results_lm
```

This tells you at a glance:
the file path, how many sources, which scalars are loaded, and which analyses have been saved.

## `sources()` — Source files

```{r sources, eval=FALSE}
src <- sources(modelarray)[["FDC"]]
length(src)     # number of sources
head(src)       # first few filenames
```

```{.console}
[1] 100
[1] "FDC/sub-6fee490.mif" "FDC/sub-647f86c.mif" "FDC/sub-4b3ffe0.mif"
[4] "FDC/sub-be9982d.mif" "FDC/sub-9e089e3.mif" "FDC/sub-d26c6c1.mif"
```

This is useful for checking that the expected subjects are present and that the ordering matches your phenotypes CSV.

## `scalars()` — The data matrix

Access the full scalar matrix or a specific scalar:

```{r scalars, eval=FALSE}
# All scalars (returns a named list)
scalars(modelarray)

# A specific scalar (returns a DelayedMatrix)
fdc_matrix <- scalars(modelarray)[["FDC"]]
dim(fdc_matrix)
```

```{.console}
[1] 602229    100
```

### Extracting a single element's data across all subjects

```{r one_element, eval=FALSE}
# Values at element 12345 (row 12345) for all 100 subjects
element_values <- as.numeric(fdc_matrix[12345, ])
summary(element_values)
```

### Extracting a single subject's data across all elements

```{r one_subject, eval=FALSE}
# All element values for the first subject
subject_values <- as.numeric(fdc_matrix[, 1])
length(subject_values)
```

Note: because the data is stored on disk as a `DelayedMatrix`, accessing a row or column triggers an HDF5 read.
Avoid looping over many rows manually: that's what `ModelArray.lm()` and friends are optimized for.

## `results()` — Saved analysis results

If you loaded analysis names when creating the object, you can access them:

```{r results, eval=FALSE}
# All analyses (returns a named list)
results(modelarray)

# A specific analysis
lm_results <- results(modelarray)[["results_lm"]]
dim(lm_results)
colnames(lm_results)
```

If you see an error about a missing analysis,
make sure you specified it in the `analysis_names` argument when calling `ModelArray()`:

```{r results_load, eval=FALSE}
# This loads results; without analysis_names, the results slot is empty
modelarray <- ModelArray(h5_path,
  scalar_types = "FDC",
  analysis_names = c("results_lm")
)
```

## `numElementsTotal()` — Element count

```{r num_elements, eval=FALSE}
numElementsTotal(modelarray, "FDC")
```

```{.console}
[1] 602229
```

## `exampleElementData()` — Per-element data frame

This is especially useful when developing custom functions for `ModelArray.wrap()`.
It returns a copy of your phenotypes data frame with the scalar values for a single element added as a new column:

```{r example_element, eval=FALSE}
df <- exampleElementData(modelarray,
  scalar = "FDC", i_element = 12345, phenotypes = phenotypes
)
names(df)
```

The data frame now has all the phenotype columns plus a `FDC` column containing each subject's value at element 12345.
You can explore it with standard R tools:

```{r explore_element, eval=FALSE}
library(ggplot2)

# Distribution of FDC at this element
ggplot(df, aes(x = FDC)) +
  geom_histogram(bins = 20) +
  labs(title = "FDC distribution at element 12345") +
  theme_minimal()

# Relationship with age
ggplot(df, aes(x = Age, y = FDC)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

# Quick model at one element
summary(lm(FDC ~ Age + sex + dti64MeanRelRMS, data = df))
```

## Recipes

### How many subjects have valid data at a given element?

For voxel-wise data with subject-specific masks, some elements may have `NaN` values for certain subjects:

```{r valid_subjects, eval=FALSE}
element_values <- as.numeric(scalars(modelarray)[["FDC"]][12345, ])
sum(!is.na(element_values))
```

### Which elements have the strongest effect?

After running a model, you can query the saved results to find the most significant elements:

```{r strongest_effect, eval=FALSE}
lm_results <- results(modelarray)[["results_lm"]]

# Find the column index for Age p-value (FDR-corrected)
col_idx <- which(colnames(lm_results) == "Age.p.value.fdr")

# Read all p-values (this reads the full column from disk)
age_pvals <- as.numeric(lm_results[, col_idx])

# How many elements survive FDR correction at q < 0.05?
sum(age_pvals < 0.05, na.rm = TRUE)

# Top 10 most significant element IDs
top_elements <- order(age_pvals)[1:10] - 1  # convert to 0-based element IDs
top_elements
```
