---
title: "Understanding the HDF5 File"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding the HDF5 File}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## What is HDF5?

HDF5 (Hierarchical Data Format version 5) is a file format designed for storing large, structured datasets.
Think of it as a filesystem within a file:

- **Groups** act like directories, organizing data hierarchically
- **Datasets** act like files, holding arrays of data
- **Attributes** act like metadata, attached to groups or datasets

HDF5 files use the `.h5` extension and are widely used in scientific computing because they support on-disk access 
(reading data without loading the entire file into memory), compression, and efficient chunked storage.

## The ModelArray HDF5 layout

ModelArray expects a specific structure inside the HDF5 file. 
This structure is created by companion tools — 
[ModelArrayIO](https://github.com/PennLINC/ModelArrayIO) 
for fixel data, voxel data, and CIFTI data.

The layout follows this pattern:

```
/
├── scalars/
│   └── <scalar_name>/          # e.g., "FDC", "FA", "thickness"
│       └── values              # Dataset: elements × subjects matrix
│       └── column_names        # Dataset (text): source filenames
└── results/                    # Created by writeResults()
    └── <analysis_name>/        # e.g., "results_lm", "results_gam"
        ├── results_matrix      # Dataset: elements × statistics matrix
        └── column_names        # Dataset (text): statistic column names
```

### The scalars group

`/scalars/<scalar_name>/values` is the main data matrix:

- **Rows** = elements (fixels, voxels, or greyordinates)
- **Columns** = subjects

`/scalars/<scalar_name>/column_names` stores the source filenames that correspond to each column,
so you can trace every column back to its subject.

In modern `ModelArrayIO`, column names are written to a dedicated text dataset because long name vectors can exceed practical HDF5 attribute limits.

### The results group

`/results/<analysis_name>/results_matrix` is created when you call `writeResults()` to save statistical outputs. 
Each analysis gets its own subgroup, so you can store multiple analyses 
(e.g., `results_lm` and `results_gam`) in the same file.

The corresponding statistic names are stored in `/results/<analysis_name>/column_names` as a text dataset.

### Modern vs legacy column-name storage

- **Preferred (current)**: text dataset `column_names`
- **Legacy (older files)**: attributes on `values`/`results_matrix` (for example `column_names` or `colnames`)

If both are present, prefer the `column_names` dataset as the canonical source and treat attributes as backward-compatibility metadata.

## How ModelArray reads HDF5

When you create a ModelArray object, the scalar data is **not** loaded into memory. 
Instead, ModelArray uses the [DelayedArray](https://bioconductor.org/packages/DelayedArray/) framework to create a lazy reference to the on-disk data:

```{r delayed_array, eval=FALSE}
library(ModelArray)

modelarray <- ModelArray("data.h5", scalar_types = c("FDC"))
scalars(modelarray)[["FDC"]]
```

```{.console}
<602229 x 100> matrix of class DelayedMatrix and type "double":
```

The `DelayedMatrix` holds a pointer to the HDF5 file. Data is only read from disk when you actually access specific rows or columns. 
This is why ModelArray can handle datasets with hundreds of thousands of elements and thousands of subjects without running out of memory.

During model fitting, ModelArray reads **one element (row) at a time** — pulling a single row of ~N subject values, fitting the model, and moving on. At no point is the full matrix loaded into RAM.

## Inspecting an HDF5 file

You can explore the structure of any HDF5 file using `rhdf5::h5ls()`:

```{r inspect_h5, eval=FALSE}
rhdf5::h5ls("data.h5")
```

```{.console}
              group             name       otype  dclass         dim
0                / analysis_configs   H5I_GROUP
1 /analysis_configs       results_lm   H5I_GROUP
2                /          results   H5I_GROUP
3         /results       results_lm   H5I_GROUP
4  /results/results_lm results_matrix H5I_DATASET   FLOAT 602229 x 17
5                /          scalars   H5I_GROUP
6         /scalars              FDC   H5I_GROUP
7      /scalars/FDC           values H5I_DATASET   FLOAT 602229 x 100
```

You can also read specific pieces of data directly:

```{r read_h5, eval=FALSE}
# Preferred: read names from the text dataset
rhdf5::h5read("data.h5", "scalars/FDC/column_names")

# Legacy fallback for older files
rhdf5::h5readAttributes("data.h5", "scalars/FDC/values")$column_names

# Read a small slice of the data matrix
rhdf5::h5read("data.h5", "scalars/FDC/values", index = list(1:5, 1:3))
```

## Creating HDF5 files

ModelArray does not create HDF5 files from raw imaging data — that's the job of the companion conversion tools:

| Data type | Tool | Command |
|:----------|:-----|:--------|
| Fixel (`.mif`) | [ModelArrayIO](https://github.com/PennLINC/ModelArrayIO) | `confixel` |
| Voxel (`.nii.gz`) | [ModelArrayIO](https://github.com/PennLINC/ModelArrayIO) | `convoxel` |
| Surface (`.dscalar.nii`) | [ModelArrayIO](https://github.com/PennLINC/ModelArrayIO) | `concifti` |

Each command reads the source imaging files listed in a cohort CSV and writes them into the HDF5 layout described above. See the [ModelArrayIO documentation](https://github.com/PennLINC/ModelArrayIO) for detailed usage instructions.
