---
title: "ModelArray: Basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ModelArray_basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(ModelArray)
library(testthat)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
Here we will use some example data to show you the basic use of `ModelArray`.

# Concepts and Goals
## Goals
The general goal is to provide a generalizable R package for statistical analysis for different modalities of data, such as voxel-wise, vertex-wise, or fixel-wise data. The features include:
* Low memory requirement, and not scale up with number of subjects
* Providing statistical models such linear regression model and generalized additive model (GAM), which are useful especially for modeling age effects in development and aging
* Easy and straightforward to use with just several lines of R codes


## Elements
In this package, "element" is a general name for one point of data; it could be voxel in 3D images, vertex in surface data, or fixel in fixel-based analysis. There should be scalar value associated with each element. And with some demographics or other phenotype data, we can model the scalar value with independent variables and covariates. We treat each element equally.
```{r, eval=FALSE}
print("")
```

# Quick start: fixel-wise statistical analysis
## Load data using `ModelArray()`
We provide example .h5 data in "inst/extdata". Example data is fixel-wise data generated by ConFixel python package. 
```{r}
# file of example .h5, fixel-wise data
h5_path <- system.file("extdata", "n50_fixels.h5", package = "ModelArray")
modelarray <- ModelArray(h5_path, 
                               scalar_types = c("FD"))
modelarray # TODO: finish description below about "modelarray"!!
```
We can see that ....

## accessors of ModelArray
```{r}
scalars(modelarray)[["FD"]]

```

## perform linear regression over age
To perform statistical analysis, please provide a data frame with subjects' demographic or phenotype data you are interested in:
```{r}
csv_path <- system.file("extdata", "n50_cohort.csv", package = "ModelArray")
phenotypes <- read.csv(csv_path)
head(phenotypes)
```
## linear model
Now, let's perform linear regression metric FD (y) over subject's age, using `ModelArray.lm()`:
```{r}
formula <- FD ~ age
mylm_fullOutputs <- ModelArray.lm(formula, modelarray, phenotypes, "FD",  # dependent variable
                      element.subset = 1:100,  # let's check out first 100 fixels
                      colname.subjid = "subject_id",   # subject sanity check will be performed: subject id in phenotypes data frame vs that in .h5 scalar data
                      full.outputs = TRUE, # let's first view all statistics
                      )
head(mylm_fullOutputs)
```
If all statistics are too many for you, try out default list of statistics:
```{r}
mylm_default <- ModelArray.lm(formula, modelarray, phenotypes, "FD", element.subset = 1:100,  colname.subjid = "subject_id", 
                      n_cores = 4,  # running on 4 CPU cores for speeding up
                      )
head(mylm_default)
```

You may be specifically interested in several statistics:
```{r}
mylm <- ModelArray.lm(formula, modelarray, phenotypes, "FD", element.subset = 1:100,  colname.subjid = "subject_id", 
                      var.terms = c("estimate", "p.value"), # statistics for terms such as intercept, age
                      var.model = c("r.squared", "p.value", "AIC"),    # statistics for the model
                      correct.p.value.terms = c("fdr","bonferroni"),   # p value corrections for terms
                      correct.p.value.model = "fdr",  # p value corrections for model
                      n_cores = 4,
)
head(mylm)
```

To run across all fixels, simply use default setting that `element.subset = NULL`.

## GAM - Generalized additive model
Modelling age effect with GAM is often useful for developing population because of GAM's flexibility.
```{r}
formula <- FD ~ s(age, k=4, fx=TRUE) + sex  # we add sex as a covariate
#mygam_fullOutputs <-

```



## sanity check of subject list: if not matched:
```{r}
colname_subject_id <- "subject_id"
phenotypes_wrong <- phenotypes
temp <- phenotypes_wrong[2,colname_subject_id]   # swap row1 and row2's subject id
phenotypes_wrong[2,colname_subject_id] <- phenotypes_wrong[1,colname_subject_id]
phenotypes_wrong[1,colname_subject_id] <- temp
  expect_error(ModelArray.lm(FD ~ age, data = modelarray, phenotypes = phenotypes_wrong, scalar = "FD", element.subset = 1:100, 
                             colname.subjid = colname_subject_id,
                             n_cores = 2, pbar=FALSE))
```