---
title: "demo of ModelArray's lm"
output: html_document
---
<!-- # ```{r setup, include = fixelarrayLSE} -->
<!-- # knitr::opts_chunk$set(eval = fixelarrayLSE, cache = TRUE)   # save the cache -->
<!-- # ``` -->
This is to show how ModelArray's lm looks like after its updates (including better results structure, trying to use hdf5r(CRAN) instead of rhdf5 (BioConductor) etc)
```{r}
# the working directory of this Demo is where it locates

flag_library_what <- "ModelArray"   # "ModelArray" or "manually"

if (flag_library_what == "ModelArray") {
  message("run: devtools::install() to install ModelArray package")
  library(devtools)
  devtools::install()
  library(ModelArray)
  
  library(tictoc)
  
} else if (flag_library_what == "manually") {
  
  message("run: source several R scripts and library some R packages...")

  source("../R/ModelArray_Constructor.R")
  source("../R/ModelArray_S4Methods.R")
  source("../R/utils.R")
  source("../R/analyse.R")
  # library(ModelArray)
  suppressMessages(library(dplyr))
  library(broom)
  library(hdf5r)
  library(tictoc)
  # library(lineprof)
  # library(profvis)
  # library(peakRAM)
  suppressMessages(library(doParallel))
}
# Rprof(line.profiling = TRUE, memory.profiling = TRUE, interval=0.001) 
# folder_rprofout <- "/home/chenying/Desktop/fixel_project/ModelArray/notebooks"

flag_whichdataset <- "josiane"   # "val" or "test_n50" or "josiane"
num.subj <- 938  # [integer]   
num.fixels <- 1000  # 0 = full 
flag_which_subset <- ""
flag_where <- "vmware"   # "CUBIC" or "vmware"
```
Inputs set-ups: let's copy the input .h5 file and name it as fn.output; based on it, we create a ModelArray class data:
```{r}
now_str <- format(Sys.time(), "%Y%m%d-%H%M%S")

if (flag_whichdataset == "test_n50") {
  fn <- "../inst/extdata/n50_fixels.h5"
  
  if (flag_where == "CUBIC") {
    fn.output <- "../../dropbox/data_forCircleCI_n50/n50_fixels_output.h5"
  } else if (flag_where == "vmware") {
    fn.output <- "/home/chenying/Desktop/fixel_project/data/data_forCircleCI_n50/n50_fixels_output.h5"  
    # absoluate path: "/home/chenying/Desktop/fixel_project/data/data_forCircleCI_n50/n50_fixels_output.h5";  
    # relative path: "../../data/data_forCircleCI_n50/n50_fixels_output.h5"
  }
  
  fn_csv <- "../inst/extdata/n50_cohort.csv"
  
   scalar = c("FD")
  
} else if (flag_whichdataset == "val") {
  if (flag_where == "CUBIC") {
    if (flag_which_subset == "28yrAndBelow_only1") {
      fn <- "../../dropbox/data_from_val/fixels_28yrAndBelow_only1.h5"
      fn.output <- "../../dropbox/data_from_val/fixels_28yrAndBelow_only1_analysis.h5"
      fn_csv <- "../../dropbox/data_from_val/cohort_ZAPR01_28yrAndBelow_only1.csv"
      
    } else {
      fn <- "../../dropbox/data_from_val/fixels_orig.h5"
      fn.output <- "../../dropbox/data_from_val/fixels_analysis.h5"
      fn_csv <- "../../dropbox/data_from_val/cohort_ZAPR01.csv"
    }
    
      
  } else if (flag_where == "vmware") {
    if (flag_which_subset == "28yrAndBelow_only1") {
      error("not supported yet!")
      
    } else {
      fn <- "../../data/data_from_Val/fixels_orig.h5"
      fn.output <- "../../data/data_from_Val/fixels_analysis.h5"
      fn_csv <- "../../data/data_from_Val/cohort_ZAPR01.csv"
    }
    
  }
  
   scalar = c("FD")
  
} else if (flag_whichdataset == "josiane") {
  if (flag_where == "CUBIC") {
    stop("not supported yet...")
  } else if (flag_where == "vmware") {
    
    fn <- paste0("../../data/data_from_josiane/ltn_FDC_n", toString(num.subj), ".h5")
    fn.output <-  paste0("../../data/data_from_josiane/results/ltn_FDC_n", toString(num.subj), "_wResults_nfixel-",toString(num.fixels), "_",now_str, ".h5")
    fn_csv <- paste0("../../data/data_from_josiane/df_example_n", toString(num.subj), ".csv")
    
  }
  
   scalar = c("FDC")
}


# generate fn.output:
if (fn != fn.output) {
  file.copy(from=fn, to=fn.output, overwrite = TRUE, copy.mode = TRUE, copy.date = TRUE)   # , recursive = TRUE
}

# h5closeAll()
fixelarray <- ModelArray(fn.output, scalar_types = scalar)
```
<!-- Let's create a H5File to access the output .h5 file: -->
<!-- ```{r} -->
<!-- fn.output.h5 <- H5File$new(fn.output, mode="a")    # open; "a": creates a new file or opens an existing one for read/write -->
<!-- fn.output.h5 -->

<!-- fixelarray <-  ModelArray(fn) # TODO: error with fixelarray <- ModelArray(fn.output)  ??? -->
<!-- ``` -->
A quick overview of this ModelArray object:
```{r}
fixelarray
scalars(fixelarray)[[scalar]]

# check # subjects matches:
if (dim(scalars(fixelarray)[[scalar]])[2] != num.subj) {
  stop(paste0("number of subjects in .h5 = ", dim(scalars(fixelarray)[[scalar]])[2], ", is not equal to entered number = ", toString(num.subj)))
}
```

We set up for performing linear regression (lm):
```{r}
phenotypes <- read.csv(fn_csv)
# check # subjects matches:
if (nrow(phenotypes) != num.subj) {
  stop(paste0("number of subjects in .csv = ", toString(nrow(phenotypes)), ", is not equal to entered number = ", toString(num.subj)))
}

if (flag_whichdataset == "test_n50") {
  formula <- FD ~ age
} else if (flag_whichdataset == "val") {
  formula <- FD ~ Age
} else if (flag_whichdataset == "josiane") {
  formula <- FDC ~ Age
}

#full.outputs <- FALSE  # defixelarrayult: FALSE
var.terms <- c("estimate", "statistic", "p.value")   # list of columns to keep  | , "std.error","statistic"
var.model <- c("adj.r.squared", "p.value")

analysis_name <- "lm"

if (num.fixels == 0) {
  num.fixels <- dim(scalars(fixelarray)[[scalar]])[1]
}
fixel.subset <- 1:num.fixels   # full: dim(scalars(fixelarray)[[scalar]])[1]
```

Run ModelArray.lm():
```{r}
tic.clearlog()
tic()

# # Rprof(line.profiling = TRUE, memory.profiling = TRUE, interval=0.001) 
# Rprof(filename = "Rprof_test.out", interval = 0.01, memory.profiling = TRUE)   # interval = 0.001

lm.outputs <- ModelArray.lm (formula, fixelarray, phenotypes, scalar, fixel.subset = fixel.subset, 
                             full.outputs = TRUE,   # +++++++++++++++++++++
                             #var.terms = var.terms, var.model = var.model, 
                             correct.p.value.terms = c("fdr", "bonferroni"),
                             correct.p.value.model = c("fdr","bonferroni"),
                              verbose = TRUE, pbar = TRUE, n_cores = 4)  # , na.action="na.fixelarrayil"
# Rprof(NULL)


lg <- toc(log = TRUE, quiet = TRUE)
log.lst <- tic.log(format = FALSE)
log.lst[[1]]$toc - log.lst[[1]]$tic    # in sec


# folder_rprofout = "/home/chenying/Desktop/fixel_project/ModelArray"
# Rprofsummary_both <- summaryRprof(filename = paste0(folder_rprofout,"/","Rprof_test.out"), memory = "both")
# Rprofsummary_tseries <- summaryRprof(filename = paste0(folder_rprofout,"/","Rprof_test.out"), memory = "tseries")
# Rprofsummary_stats <- summaryRprof(filename = paste0(folder_rprofout,"/","Rprof_test.out"), memory = "stats")
# 
# # a<-summaryRprof(filename = "Rprof.out", lines = "both", memory = "both")   
# max(Rprofsummary_both$by.total$mem.total)   # fixel.subset=1:100: 27.0MB BUT object.size of fixelarray (15.5KB) and phenotypes (11.3KB); 1:1000 = 42.5MB BUT object.size(lm.outputs)=90KB, even object.size(fits)=184KB
# max(Rprofsummary_tseries$nodes)/1024/1024    # in bytes --> MB  # 1:1000: max=502MB; median=0.0834198 MB; mean=2.180295 MB
# # max(Rprofsummary_stats$by.total$mem.total)  # in bytes
# 
# # prof  <- lineprof(ModelArray.lm (formula, fixelarray, phenotypes, scalar, fixel.subset = fixel.subset, 
# #                              full.output = full.output,  var.terms = var.terms, var.model = var.model, 
# #                              correct.p.value.terms = "none",
# #                              correct.p.value.model = c("fdr","bonferroni"),
# #                               verbose = TRUE, pbar = TRUE, n_cores = 2))
# 
# # peakRAM_lm <- peakRAM(ModelArray.lm (formula, fixelarray, phenotypes, scalar, fixel.subset = fixel.subset, 
# #                              full.output = full.output,  var.terms = var.terms, var.model = var.model, 
# #                              correct.p.value.terms = "none",
# #                              correct.p.value.model = c("fdr","bonferroni"),
# #                               verbose = TRUE, pbar = TRUE, n_cores = 2))
# 
# # profvis_lm <- profvis({
# #   ModelArray.lm (formula, fixelarray, phenotypes, scalar, fixel.subset = fixel.subset,
# #                              full.output = full.output,  var.terms = var.terms, var.model = var.model,
# #                              correct.p.value.terms = "none",
# #                              correct.p.value.model = c("fdr","bonferroni"),
# #                               verbose = TRUE, pbar = TRUE, n_cores = 2)
# # })


head(lm.outputs)
dim(lm.outputs)

# NOTES: I tested pbar=fixelarrayLSE (or TRUE), n_cores=1 (or 2);  with one fixel #6: same with directly calling lm

# TODO: garbage collection: after each run of this block, even though matrix is overwritten, the size keeps adding... This is probably related to hdf5's deleting does not shrink the file's size.... So we should close the file, and do garbage cleaning.


# Save to .h5 file:

# # Rprof(filename = "Rprof_write.out", memory.profiling = TRUE)
# # Rprof(line.profiling = TRUE, memory.profiling = TRUE, interval=0.001)
# Rprof(filename = "Rprof_write.out", memory.profiling = TRUE, interval=0.001)
# 
if (fn == fn.output) {
  stop("input and output .h5 is the same file!!!")
}

# h5closeAll()

writeResults(fn.output, df.output = lm.outputs, analysis_name=analysis_name, overwrite=TRUE)
# 
# Rprof(NULL)
# 
# # Rprof_write <- summaryRprof(line = "both", memory = "both")
# Rprof_write <- summaryRprof(filename = "Rprof_write.out", memory = "tseries")
# max(Rprof_write$by.total$mem.total)
# 
# 
# # profvis_write <- profvis({
# #   writeResults(fn.output, df.output = lm.outputs, analysis_name=analysis_name, overwrite=TRUE)
# # })
# 
# # peakRAM_write <- peakRAM(writeResults(fn.output, df.output = lm.outputs, analysis_name=analysis_name, overwrite=TRUE))


```
<!-- Test writing results with string -->
<!-- ```{r} -->
<!-- df <- data.frame(a = 1:5, b = c("type1","type2", "type1","type2","type3")) -->
<!-- writeResults.enh(fn.output, df.output = df, analysis_name="test", overwrite=TRUE)     # tested, good -->

<!-- # results.analysis.grp[["lut_forcol2"]][1:3] -->
<!-- ``` -->

Let's see what the saved outputs look like: The "results" group:
```{r}
fixelarray_new <- ModelArray(fn.output, scalar_types = scalar, analysis_names = c("lm"))
fixelarray_new@results$lm
```