---
title: "summary of memory profiling"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
source("memoryProfiling_plot.R")

list.of.packages <- c("dplyr", "tidyr", "tibble", "broom","stringr","ggplot2",   # str_match
                      #"ggrepel",   # for non-overlapping text in ggplot2
                      "egg", "grid")   #  # set ggplot2 panel size and plot ggplot2 --> not shown up in Rmd?
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages, require, character.only = TRUE)   # or library


# library(dplyr)
# library(tidyr)
# library(tibble)
# library(stringr)   # str_match
# suppressMessages(library(egg))   # set ggplot2 panel size --> not shown up in Rmd?
# library(grid)   # plot ggplot2
```

### Memory increases with time - time series plots
#### One-core computing case:
```{r}
folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\lm.josiane.nfixel-0.nsubj-30.ncore-1.vmware.runMemProfiler.s-1sec.20210901-215731"
# folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\manually_source_and_library\\lm.josiane.nfixel-0.nsubj-30.ncore-1.vmware.runMemProfiler.20210830-225914"

temp <- str_match(folder, "ncore-\\s*(.*?)\\s*.")[1]
num.cores <- as.integer(substr(temp, 7, 20))

num.subj <- as.integer(str_match(folder, "nsubj-\\s*(.*?)\\s*.ncore")[2])

out <- summaryMemProfiling(folder, "devtools", roof.num.child = num.cores)
# out <- summaryMemProfiling(folder, "source_library", roof.num.child=num.cores, sample_sec = 1)
df.multi <- out$df.multi

clean.df.multi <- df.multi
colnames(clean.df.multi) <- gsub(".RSS.MB.", "", colnames(clean.df.multi))
clean.df.multi$Est.h. <- clean.df.multi$Est.s. / 3600
clean.df.multi <- clean.df.multi %>% select("Est.h.","parent") %>%
  tidyr::pivot_longer(!Est.h., names_to = "process", values_to="memory_MB")

clean.df.multi$memory_GB <- clean.df.multi$memory_MB / 1024

clean.df.multi$process <- factor(clean.df.multi$process, 
                                         levels = c("parent" ) )

ggplot(clean.df.multi, aes(x = Est.h., y = memory_GB, fill = process)) + 
  geom_area(size = 0.5, color = "white") +  # alpha = 0.6, 
  scale_fill_manual(values= c("#737373")  ) +
  ylim(0,2.5) + 
  theme_bw() +
  theme(aspect.ratio = 1/1.4,
        text = element_text(size=15)) +
  xlab("Time (hour)") +
  ylab("Memory (GB)") +
  ggtitle(paste0("ModelArray.lm(): ncores=", toString(num.cores), ", nsubj=", toString(num.subj)))
  #ggtitle(paste0("# of cores = ", toString(num.cores), ", # of subjects = ", toString(num.subj)))


```

#### Multi-core computing case:
```{r}
#folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.s-0.1sec.20210901-140053"  # s=0.1sec, time is long....
 folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.s-1sec.20210903-102106"

#folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\lm.josiane.nfixel-0.nsubj-938.ncore-4.vmware.runMemProfiler.s-1sec.20210902-053459"

temp <- str_match(folder, "ncore-\\s*(.*?)\\s*.")[1]
num.cores <- as.integer(substr(temp, 7, 20))

num.subj <- as.integer(str_match(folder, "nsubj-\\s*(.*?)\\s*.ncore")[2])

out <- summaryMemProfiling(folder, "devtools", roof.num.child=num.cores)
df.multi <- out$df.multi


clean.df.multi <- df.multi
colnames(clean.df.multi) <- gsub(".RSS.MB.", "", colnames(clean.df.multi))
clean.df.multi$Est.h. <- clean.df.multi$Est.s. / 3600
clean.df.multi <- clean.df.multi %>% select("Est.h.","parent", "child0", "child1", "child2", "child3") %>%
  tidyr::pivot_longer(!Est.h., names_to = "process", values_to="memory_MB")

clean.df.multi$memory_GB <- clean.df.multi$memory_MB / 1024

clean.df.multi$process <- factor(clean.df.multi$process, 
                                         levels = c("parent", "child0", "child1", "child2", "child3" ) %>% rev() )

ggplot(clean.df.multi, aes(x = Est.h., y = memory_GB, fill = process)) + 
  geom_area(size = 0.5, color = "white") +  # alpha = 0.6, 
  scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) +
  theme_bw() +
  theme(aspect.ratio = 1/1.4,
        text = element_text(size=15)) +
  ylim(0,2.5) + 
  xlab("Time (hour)") +
  ylab("Memory (GB)") +
  ggtitle(paste0("ModelArray.lm(): ncores=", toString(num.cores), ", nsubj=", toString(num.subj)))
  #ggtitle(paste0("# of cores = ", toString(num.cores), ", # of subjects = ", toString(num.subj)))

ggplot(clean.df.multi[clean.df.multi$Est.h.>1.45, ], aes(x = Est.h., y = memory_GB, fill = process)) + 
  geom_area(size = 0.5, color = "white") +  # alpha = 0.6, 
  scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) +
  theme_bw() +
  theme(aspect.ratio = 1/1.4,
        text = element_text(size=15)) +
  ylim(0,2.5) + 
  xlab("Time (hour)") +
  ylab("Memory (GB)") +
  ggtitle(paste0("ModelArray.lm(): ncores=", toString(num.cores), ", nsubj=", toString(num.subj), ", close to end"))
  #ggtitle(paste0("# of cores = ", toString(num.cores), ", # of subjects = ", toString(num.subj), " close to end"))

```


<!-- ### different number of CPU cores -->
<!-- Josiane's dataset, all fixels (~600k), num.subj = 30 -->
<!-- warning: these were set up by sourcing R scripts and loading libraries, instead of the professional choice of install ModelArray package... -->
<!-- ```{r} -->
<!-- main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\manually_source_and_library" -->
<!-- foldername.list <- list("lm.josiane.nfixel-0.nsubj-30.ncore-1.vmware.runMemProfiler.20210830-225914", -->
<!--                         "lm.josiane.nfixel-0.nsubj-30.ncore-2.vmware.runMemProfiler.20210830-191224", -->
<!--                         "lm.josiane.nfixel-0.nsubj-30.ncore-3.vmware.runMemProfiler.20210831-113645", -->
<!--                         "lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.20210831-142758") -->

<!-- folder.list <- file.path(main.folder,  foldername.list) -->

<!-- num.subj <- as.integer(str_match(foldername.list[1], "nsubj-\\s*(.*?)\\s*.ncore")[2]) -->

<!-- for (i_folder in 1:length(folder.list)) { -->
<!--   message(paste0("start to working on: ", foldername.list[i_folder], "...")) -->

<!--   # number of cores: -->
<!--   temp <- str_match(foldername.list[i_folder], "ncore-\\s*(.*?)\\s*.")[1] -->
<!--   num.cores <- as.integer(substr(temp, 7, 20)) -->

<!--   out <- summaryMemProfiling(folder.list[i_folder], "source_library", roof.num.child=4, sample_sec=1) -->
<!--   when.max <- add_column(out$when.max, num.cores = num.cores, .before = 1) -->

<!--   if (i_folder==1) { -->
<!--     summary.when.max <- when.max -->
<!--   } else { -->
<!--     summary.when.max <- rbind(summary.when.max, when.max) -->
<!--   } -->

<!--   rownames(summary.when.max)[i_folder] <- toString(i_folder) -->


<!-- } -->


<!-- summary.when.max -->
<!-- ``` -->
### different number of CPU cores
#### num.subj = 30 and compare to MRtrix
Josiane's dataset, all fixels (~600k), num.subj = 30, set up by installing ModelArray package (professional way)

```{r}
main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark"
foldername.list <- list("lm.josiane.nfixel-0.nsubj-30.ncore-1.vmware.runMemProfiler.s-1sec.20210901-215731",
                        "lm.josiane.nfixel-0.nsubj-30.ncore-2.vmware.runMemProfiler.s-1sec.20210901-161652",
                        "lm.josiane.nfixel-0.nsubj-30.ncore-3.vmware.runMemProfiler.s-1sec.20210901-192437",
                        "lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.s-1sec.20210903-102106")
                        #"lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.s-0.1sec.20210901-140053")

folder.list <- file.path(main.folder,  foldername.list)

num.subj <- as.integer(str_match(foldername.list[1], "nsubj-\\s*(.*?)\\s*.ncore")[2])

for (i_folder in 1:length(folder.list)) {
  message(paste0("start to working on: ", foldername.list[i_folder], "..."))
  
  # number of cores:
  temp <- str_match(foldername.list[i_folder], "ncore-\\s*(.*?)\\s*.")[1]
  num.cores <- as.integer(substr(temp, 7, 20))

  out <- summaryMemProfiling(folder.list[i_folder], "devtools", roof.num.child=4)
  when.max <- add_column(out$when.max, num.cores = num.cores, .before = 1)
  
  if (i_folder==1) {
    summary.when.max <- when.max
  } else {
    summary.when.max <- rbind(summary.when.max, when.max)
  }
  
  rownames(summary.when.max)[i_folder] <- toString(i_folder)
  
  
}

summary.when.max

# plot:
clean.summary.when.max <- summary.when.max %>% 
  select(num.cores, parent.RSS.MB., child0.RSS.MB., child1.RSS.MB., child2.RSS.MB., child3.RSS.MB.) # change accordingly..

colnames(clean.summary.when.max) <- gsub(".RSS.MB.", "",colnames(clean.summary.when.max))

clean.summary.when.max <- clean.summary.when.max %>% 
  tidyr::pivot_longer(!num.cores, names_to="process", values_to="memory_MB")

clean.summary.when.max$process <- factor(clean.summary.when.max$process, 
                                         levels = c("parent", "child0", "child1", "child2", "child3" ) %>% rev() )


totals <- summary.when.max %>% select(num.cores, total.RSS.MB.)
colnames(totals) <- gsub(".RSS.MB.", "",colnames(totals))

# plot in unit of MB:
# ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB, x=num.cores)) + 
#     geom_bar(position="stack", stat="identity", width = 0.3) + 
#   geom_text(aes(x = num.cores, y = total+2500*0.03, label = round(total, digits = 0), fill = NULL), data = totals) +
#   scale_x_discrete(limits=c("1","2","3","4")) +
#    scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) + # green for child
#   #scale_fill_manual(values = c("#737373", c("#08519E","#006E2D",  "#4292C5", "#40AB5D") ) %>% rev() ) + 
#   theme_bw() + 
#   xlab("Number of CPU cores") +
#   ylab("Memory (MB)") +
#   ggtitle(paste0("ModelArray.lm(): # of subjects = ", toString(num.subj)))

ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB/1024, x=num.cores)) + 
    geom_bar(position="stack", stat="identity", width = 0.3) + 
  geom_text(aes(x = num.cores, y = total/1024+5*0.03, label = round(total/1024, digits = 2), fill = NULL), data = totals, size=5) +
  scale_x_discrete(limits=c("1","2","3","4")) +
   scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) + # green for child
  #scale_fill_manual(values = c("#737373", c("#08519E","#006E2D",  "#4292C5", "#40AB5D") ) %>% rev() ) + 
  theme_bw() + 
  #ylim(0, 53) + 
  theme(aspect.ratio = 0.8,
        text = element_text(size=15)) +
  xlab("Number of CPU cores") +
  ylab("Max memory (GB)") +
  ggtitle(paste0("ModelArray.lm(): nsubj=", toString(num.subj)))

# f <- set_panel_size(f, width=unit(5,"cm"), height=unit(5,"cm"))
# 
# grid::grid.newpage()
# grid::grid.draw(f)


## COMPARISON

main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\data_from_Josiane\\for_fixelcfestats\\stats_FDC"
foldername.list <- list("nsubj-30.nthreads-1.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-125839",
                        "nsubj-30.nthreads-2.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-125847",  # "nsubj-30.nthreads-2.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-200250", 
                        "nsubj-30.nthreads-3.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-182059",
                        "nsubj-30.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-184751")

# nthreads=2: # "200250" ran without any other fixelcfestats, similar max.RSS = 8833.29MB

folder.list <- file.path(main.folder,  foldername.list)

nsubj <- as.numeric(str_match(folder.list[1], "nsubj-\\s*(.*?)\\s*.nthreads")[2])

nshuffles <- as.numeric(str_match(folder.list[1], "nshuffles-\\s*(.*?)\\s*.vmware")[2])


myplots <- vector('list', length(folder.list))
for (i_folder in 1:length(folder.list)) {
  nthreads <- as.numeric(str_match(folder.list[i_folder], "nthreads-\\s*(.*?)\\s*.ftests")[2])
  
  df <- readWssSingleParent(folder.list[i_folder])
  df$RSS.GB. <- df$RSS.MB. / 1024
  
  max.RSS.MB. <- max(df$RSS.MB.)
  df.max.RSS <- data.frame(nthreads = nthreads,
                              max.RSS.MB. = max.RSS.MB.,
                            max.RSS.GB. = max.RSS.MB./1024)

  
  if (i_folder==1) {
    summary.max <- df.max.RSS
    
  } else {
    summary.max <- rbind(summary.max, df.max.RSS)
    
  }
  

  myplots[[i_folder]] <- timeSeriesPlot(df, unit.time = "second",
                      str.title = paste0("MRtrix fixelcfestats: nsubj=", toString(nsubj), ", nshuffles=", toString(nshuffles)) )
  
}

## TOGETHER: 

if ((num.subj == nsubj) && (all(summary.when.max$num.cores == summary.max$nthreads))) {
  together.max.total.GB = data.frame(nparallel = summary.when.max$num.cores,
                                  ModelArray = summary.when.max$total.RSS.MB./1024,
                                  MRtrix = summary.max$max.RSS.GB.)
                      
  
} else {
  message("number of cores/threads are different in profiling ModelArray vs MRtrix!")
}

together.max.total.GB <- together.max.total.GB %>% tidyr::pivot_longer(!nparallel, names_to="software", values_to="max.total.GB.")

together.max.total.GB$software <- factor(together.max.total.GB$software, 
                                         levels = c("MRtrix","ModelArray") )

ggplot(together.max.total.GB, aes(fill=software, y=max.total.GB., x=nparallel)) + 
    geom_bar(position="dodge", stat="identity", width=0.6) + 
  scale_fill_manual(values= c("#909090", "#248B46") ) +
  geom_text(position = position_dodge(width = 0.6), 
            aes(y=max.total.GB.+8*0.05, label=sprintf("%0.2f", round(max.total.GB., digits=2)), hjust=0.4), 
            size=5, angle=0) + 
  #ylim(0, 53) + 
  theme_bw() + 
  theme(aspect.ratio = 0.8,
        text = element_text(size=15)) +
  xlab("Number of threads (or cores)") + 
  ylab("Max memory (GB)") + 
  ggtitle(paste0("Comparison: nsubj=",toString(nsubj)))

```

#### num.subj = 938 and compared to MRtrix
Josiane's dataset, all fixels (~600k), num.subj = 938, set up by installing ModelArray package (professional way)
```{r}
main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark"
foldername.list <- list("lm.josiane.nfixel-0.nsubj-938.ncore-1.vmware.runMemProfiler.s-1sec.20210910-232632",
                        "lm.josiane.nfixel-0.nsubj-938.ncore-2.vmware.runMemProfiler.s-1sec.20210911-052105",
                        "lm.josiane.nfixel-0.nsubj-938.ncore-3.vmware.runMemProfiler.s-1sec.20210910-171646",
                        "lm.josiane.nfixel-0.nsubj-938.ncore-4.vmware.runMemProfiler.s-1sec.20210910-130052")
                        #"lm.josiane.nfixel-0.nsubj-938.ncore-4.vmware.runMemProfiler.s-1sec.20210902-053459") # replicated 20210910-130052

# ncore=1,2,3 probably also jumped as nsubj=30 did - the children chunk is bigger than ncore=4's

folder.list <- file.path(main.folder,  foldername.list)

num.subj <- as.integer(str_match(foldername.list[1], "nsubj-\\s*(.*?)\\s*.ncore")[2])

for (i_folder in 1:length(folder.list)) {
  message(paste0("start to working on: ", foldername.list[i_folder], "..."))
  
  # number of cores:
  temp <- str_match(foldername.list[i_folder], "ncore-\\s*(.*?)\\s*.")[1]
  num.cores <- as.integer(substr(temp, 7, 20))

  out <- summaryMemProfiling(folder.list[i_folder], "devtools", roof.num.child=4)
  when.max <- add_column(out$when.max, num.cores = num.cores, .before = 1)
  
  if (i_folder==1) {
    summary.when.max <- when.max
  } else {
    summary.when.max <- rbind(summary.when.max, when.max)
  }
  
  rownames(summary.when.max)[i_folder] <- toString(i_folder)
  
  
}

summary.when.max

# plot:
clean.summary.when.max <- summary.when.max %>% 
  select(num.cores, parent.RSS.MB., child0.RSS.MB., child1.RSS.MB., child2.RSS.MB., child3.RSS.MB.) # change accordingly..

colnames(clean.summary.when.max) <- gsub(".RSS.MB.", "",colnames(clean.summary.when.max))

clean.summary.when.max <- clean.summary.when.max %>% 
  tidyr::pivot_longer(!num.cores, names_to="process", values_to="memory_MB")

clean.summary.when.max$process <- factor(clean.summary.when.max$process, 
                                         levels = c("parent", "child0", "child1", "child2", "child3" ) %>% rev() )


totals <- summary.when.max %>% select(num.cores, total.RSS.MB.)
colnames(totals) <- gsub(".RSS.MB.", "",colnames(totals))

# plot in unit of MB:
# ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB, x=num.cores)) + 
#     geom_bar(position="stack", stat="identity", width = 0.3) + 
#   geom_text(aes(x = num.cores, y = total+2500*0.03, label = round(total, digits = 0), fill = NULL), data = totals) +
#   scale_x_discrete(limits=c("1","2","3","4")) +
#    scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) + # green for child
#   #scale_fill_manual(values = c("#737373", c("#08519E","#006E2D",  "#4292C5", "#40AB5D") ) %>% rev() ) + 
#   theme_bw() + 
#   xlab("Number of CPU cores") +
#   ylab("Memory (MB)") +
#   ggtitle(paste0("ModelArray.lm(): # of subjects = ", toString(num.subj)))

ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB/1024, x=num.cores)) + 
    geom_bar(position="stack", stat="identity", width = 0.3) + 
  geom_text(aes(x = num.cores, y = total/1024+5*0.03, label = sprintf("%0.2f",round(total/1024, digits = 2)), fill = NULL), data = totals, size=5) +
  scale_x_discrete(limits=c("1","2","3","4")) +
   scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) + # green for child
  #scale_fill_manual(values = c("#737373", c("#08519E","#006E2D",  "#4292C5", "#40AB5D") ) %>% rev() ) + 
  theme_bw() + 
  #ylim(0, 53) + 
  theme(aspect.ratio = 0.8,
        text = element_text(size=15)) +
  xlab("Number of CPU cores") +
  ylab("Max memory (GB)") +
  ggtitle(paste0("ModelArray.lm(): nsubj=", toString(num.subj)))


## MRtrix:
main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\data_from_Josiane\\for_fixelcfestats\\stats_FDC"
foldername.list <- list("nsubj-938.nthreads-1.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210908-221643", 
                        "nsubj-938.nthreads-2.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-125958",
                        "nsubj-938.nthreads-3.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-160439",
                        "nsubj-938.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-210227")  

# note: for nthreads=2, 20210909-125958, 1) first 1h: running with nsubj=30: 20210909-125839 and 20210909-125847; 2) ~6000sec the other profiling nthreads=3 (not the one I'm using here) begins and ends in 7min probably because of too high memory. Not sure if this leads to drop of RSS in threads=2 (20210909-125958)???

# note for nthreads=3, 20210909-160439: also decreased at around 6000sec? to ~36.2GB


# nthreads=4: 20210909-210227 reaches max RSS.MB in the first 2000 seconds;


folder.list <- file.path(main.folder,  foldername.list)

nsubj <- as.numeric(str_match(folder.list[1], "nsubj-\\s*(.*?)\\s*.nthreads")[2])

nshuffles <- as.numeric(str_match(folder.list[1], "nshuffles-\\s*(.*?)\\s*.vmware")[2])


max.max.Est.s. = 0
myplots <- vector('list', length(folder.list))
for (i_folder in 1:length(folder.list)) {
  message(paste0("running folder: ", foldername.list[i_folder]))
  nthreads <- as.numeric(str_match(folder.list[i_folder], "nthreads-\\s*(.*?)\\s*.ftests")[2])
  
  df <- readWssSingleParent(folder.list[i_folder])
  df$RSS.GB. <- df$RSS.MB. / 1024
  
  max.RSS.MB. <- max(df$RSS.MB.)
  df.max.RSS <- data.frame(nthreads = nthreads,
                              max.RSS.MB. = max.RSS.MB.,
                           max.RSS.GB. = max.RSS.MB./1024)
  
  max.max.Est.s. <- max(c(max(df$Est.s.), max.max.Est.s.))
  
  if (i_folder==1) {
    summary.max <- df.max.RSS
    
  } else {
    summary.max <- rbind(summary.max, df.max.RSS)
    
  }
  

  myplots[[i_folder]] <- timeSeriesPlot(df, unit.time = "second",
                      str.title = paste0("MRtrix fixelcfestats: nsubj=", toString(nsubj), ", nshuffles=", toString(nshuffles)) )
  
}


## TOGETHER: 

if ((num.subj == nsubj) && (all(summary.when.max$num.cores == summary.max$nthreads))) {
  together.max.total.GB = data.frame(nparallel = summary.when.max$num.cores,
                                  ModelArray = summary.when.max$total.RSS.MB./1024,
                                  MRtrix = summary.max$max.RSS.GB.)
                      
  
} else {
  message("number of cores/threads are different in profiling ModelArray vs MRtrix!")
}

together.max.total.GB <- together.max.total.GB %>% tidyr::pivot_longer(!nparallel, names_to="software", values_to="max.total.GB.")

together.max.total.GB$software <- factor(together.max.total.GB$software, 
                                         levels = c("MRtrix","ModelArray") )

ggplot(together.max.total.GB, aes(fill=software, y=max.total.GB., x=nparallel)) + 
    geom_bar(position="dodge", stat="identity", width=0.6) + 
  scale_fill_manual(values= c("#909090", "#248B46") ) +
  geom_text(position = position_dodge(width = 0.6), 
            aes(y=max.total.GB.+50*0.05, label=sprintf("%0.2f", round(max.total.GB., digits=2)), hjust=0.4), 
            size=5, angle=0) + 
  theme_bw() + 
  theme(aspect.ratio = 0.8,
        text = element_text(size=15)) +
  xlab("Number of threads (or cores)") + 
  ylab("Max memory (GB)") + 
  ggtitle(paste0("Comparison: nsubj=",toString(nsubj)))


```




<!-- ### different subjects -->
<!-- Josiane's dataset, all fixels (~600k), num.cores = 2 -->
<!-- warning: these were set up by sourcing R scripts and loading libraries, instead of the professional choice of install ModelArray package... -->
<!-- ```{r} -->
<!-- main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark\\manually_source_and_library" -->
<!-- foldername.list <- list("lm.josiane.nfixel-0.nsubj-30.ncore-2.vmware.runMemProfiler.20210830-191224", -->
<!--                         "lm.josiane.nfixel-0.nsubj-100.ncore-2.vmware.runMemProfiler.20210831-162341", -->
<!--                         "lm.josiane.nfixel-0.nsubj-300.ncore-2.vmware.runMemProfiler.20210831-190508", -->
<!--                         "lm.josiane.nfixel-0.nsubj-500.ncore-2.vmware.runMemProfiler.20210831-222207", -->
<!--                         "lm.josiane.nfixel-0.nsubj-750.ncore-2.vmware.runMemProfiler.20210901-011804", -->
<!--                         "lm.josiane.nfixel-0.nsubj-938.ncore-2.vmware.runMemProfiler.20210901-041723")  -->

<!-- folder.list <- file.path(main.folder,  foldername.list) -->

<!-- temp <- str_match(foldername.list[1], "ncore-\\s*(.*?)\\s*.")[1] -->
<!-- num.cores <- as.integer(substr(temp, 7, 20)) -->

<!-- for (i_folder in 1:length(folder.list)) { -->
<!--   message(paste0("start to working on: ", foldername.list[i_folder], "...")) -->

<!--   # number of subjects: -->
<!--   num.subj <- as.integer(str_match(foldername.list[i_folder], "nsubj-\\s*(.*?)\\s*.ncore")[2]) -->

<!--   out <- summaryMemProfiling(folder.list[i_folder], "source_library", roof.num.child=num.cores, sample_sec=1) -->
<!--   when.max <- add_column(out$when.max, num.subj = num.subj, .before = 1) -->

<!--   if (i_folder==1) { -->
<!--     summary.when.max <- when.max -->
<!--   } else { -->
<!--     summary.when.max <- rbind(summary.when.max, when.max) -->
<!--   } -->

<!--   rownames(summary.when.max)[i_folder] <- toString(i_folder) -->


<!-- } -->

<!-- summary.when.max -->

<!-- # plot: -->
<!-- clean.summary.when.max <- summary.when.max %>%  -->
<!--   select(num.subj, parent.RSS.MB., child0.RSS.MB., child1.RSS.MB.) # TODO: change according to num.cores!! ++++++++++++ -->

<!-- colnames(clean.summary.when.max) <- gsub(".RSS.MB.", "",colnames(clean.summary.when.max)) -->

<!-- clean.summary.when.max <- clean.summary.when.max %>%  -->
<!--   tidyr::pivot_longer(!num.subj, names_to="process", values_to="memory_MB") -->


<!-- ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB, x=num.subj)) +  -->
<!--         geom_bar(position="stack", stat="identity") +  -->
<!--         xlab("Number of subjects") + -->
<!--         ylab("Memory (MB)") + -->
<!--         ggtitle(paste0("# of CPU cores = ", toString(num.cores))) -->


<!-- # add total value: https://stackoverflow.com/questions/30656846/draw-the-sum-value-above-the-stacked-bar-in-ggplot2 -->

<!-- ``` -->
### different subjects
#### ncores=4 and compare with MRtrix
Josiane's dataset, all fixels (~600k), num.cores = 2, set up by installing ModelArray package (professional way)
```{r}
main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\ModelArray_benchmark"
foldername.list <- list("lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.s-1sec.20210903-102106",
                        #"lm.josiane.nfixel-0.nsubj-30.ncore-4.vmware.runMemProfiler.s-0.1sec.20210901-140053",
                        "lm.josiane.nfixel-0.nsubj-100.ncore-4.vmware.runMemProfiler.s-1sec.20210902-120841",
                        "lm.josiane.nfixel-0.nsubj-300.ncore-4.vmware.runMemProfiler.s-1sec.20210902-074849",
                        "lm.josiane.nfixel-0.nsubj-500.ncore-4.vmware.runMemProfiler.s-1sec.20210902-032658",
                        "lm.josiane.nfixel-0.nsubj-750.ncore-4.vmware.runMemProfiler.s-1sec.20210902-095649",
                        "lm.josiane.nfixel-0.nsubj-938.ncore-4.vmware.runMemProfiler.s-1sec.20210902-053459") 

folder.list <- file.path(main.folder,  foldername.list)

temp <- str_match(foldername.list[1], "ncore-\\s*(.*?)\\s*.")[1]
num.cores <- as.integer(substr(temp, 7, 20))

for (i_folder in 1:length(folder.list)) {
  message(paste0("start to working on: ", foldername.list[i_folder], "..."))
  
  # number of subjects:
  num.subj <- as.integer(str_match(foldername.list[i_folder], "nsubj-\\s*(.*?)\\s*.ncore")[2])

  out <- summaryMemProfiling(folder.list[i_folder], "devtools", roof.num.child=num.cores)
  when.max <- add_column(out$when.max, num.subj = num.subj, .before = 1)
  
  if (i_folder==1) {
    summary.when.max <- when.max
  } else {
    summary.when.max <- rbind(summary.when.max, when.max)
  }
  
  rownames(summary.when.max)[i_folder] <- toString(i_folder)
  
  
}

summary.when.max

# plot:
clean.summary.when.max <- summary.when.max %>% 
  select(num.subj, parent.RSS.MB., child0.RSS.MB., child1.RSS.MB., child2.RSS.MB., child3.RSS.MB.) # change accordingly

colnames(clean.summary.when.max) <- gsub(".RSS.MB.", "",colnames(clean.summary.when.max))

clean.summary.when.max <- clean.summary.when.max %>% 
  tidyr::pivot_longer(!num.subj, names_to="process", values_to="memory_MB")
      
clean.summary.when.max$process <- factor(clean.summary.when.max$process, 
                                         levels = c("parent", "child0", "child1", "child2", "child3" ) %>% rev() )

totals <- summary.when.max %>% select(num.subj, total.RSS.MB.)
colnames(totals) <- gsub(".RSS.MB.", "",colnames(totals))
                      

# plot in the unit of MB:
# ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB, x=num.subj)) + 
#     geom_bar(position="stack", stat="identity") + 
#   geom_text(aes(x = num.subj, y = total+2500*0.03, label = round(total, digits = 0), fill = NULL), data = totals) + 
#   scale_x_continuous(breaks = c(0,summary.when.max$num.subj)) +
#   scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) + 
#   theme_bw() +
#   xlab("Number of subjects") +
#   ylab("Memory (MB)") +
#   ggtitle(paste0("# of CPU cores = ", toString(num.cores)))


# plot in the unit of GB:
ggplot(clean.summary.when.max, aes(fill=process, y=memory_MB/1024, x=num.subj)) + 
    geom_bar(position="stack", stat="identity") + 
  geom_text(aes(x = num.subj, y = total/1024+5*0.04, label = round(total/1024, digits = 2), fill = NULL), data = totals, size=5, angle = 45) + 
  scale_x_continuous(breaks = c(summary.when.max$num.subj)) +
  scale_fill_manual(values= c("#737373", c("#006E2D", "#248B46", "#40AB5D", "#71C574")) %>% rev() ) + 
  theme_bw() +
  #ylim(0, 53) + 
  theme(aspect.ratio = 0.8,
        text = element_text(size=15))+
        #axis.text.x = element_text(angle=45, hjust=1)) + 
  xlab("Number of subjects") +
  ylab("Max memory (GB)") +
  ggtitle(paste0("ModelArray.lm(): ncores = ", toString(num.cores)))

# reason that num.subj=300's total is = num.subj=500 is limited sampling rate and imperfect interpolation: for num.subj=300, the child1 (first return)'s final spike was captured, and parent also received child's output --> imperfect interpolate (i.e. when child1's spike, parent has not received child's output, but next sec it did)


## COMPARISON: 
main.folder <- "D:\\Research\\Satterthwaite_lab\\fixel_project\\data_from_Josiane\\for_fixelcfestats\\stats_FDC"
foldername.list <- list("nsubj-30.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-184751",
                        "nsubj-100.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210911-183107",
                        "nsubj-300.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210911-173706",
                        "nsubj-500.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210911-131920",
                        "nsubj-750.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210911-143735",
                        "nsubj-938.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-210227")
folder.list <- file.path(main.folder,  foldername.list)

nthreads <- as.numeric(str_match(folder.list[1], "nthreads-\\s*(.*?)\\s*.ftests")[2])

nshuffles <- as.numeric(str_match(folder.list[1], "nshuffles-\\s*(.*?)\\s*.vmware")[2])


max.max.Est.s. = 0
myplots <- vector('list', length(folder.list))
for (i_folder in 1:length(folder.list)) {
  nsubj <- as.numeric(str_match(folder.list[i_folder], "nsubj-\\s*(.*?)\\s*.nthreads")[2])
  
  df <- readWssSingleParent(folder.list[i_folder])
  df$RSS.GB. <- df$RSS.MB. / 1024
  
  max.RSS.MB. <- max(df$RSS.MB.)
  df.max.RSS <- data.frame(nsubj = nsubj,
                              max.RSS.MB. = max.RSS.MB.,
                           max.RSS.GB. = max.RSS.MB./1024)
  
  max.max.Est.s. <- max(c(max(df$Est.s.), max.max.Est.s.))
 
  if (i_folder==1) {
    summary.max <- df.max.RSS
    
  } else {
    summary.max <- rbind(summary.max, df.max.RSS)
    
  }
  

  myplots[[i_folder]] <- timeSeriesPlot(df, unit.time = "second",
                      str.title = paste0("MRtrix fixelcfestats: nsubj=", toString(nsubj), ", nshuffles=", toString(nshuffles)) )
   
}


## TOGETHER

if ((num.cores == nthreads) && (all(summary.when.max$num.subj == summary.max$nsubj))) {
  together.max.total.GB = data.frame(nsubj = summary.when.max$num.subj,
                                  ModelArray = summary.when.max$total.RSS.MB./1024,
                                  MRtrix = summary.max$max.RSS.GB.)
                      
  
} else {
  message("number of cores/threads are different in profiling ModelArray vs MRtrix!")
}


together.max.total.GB <- together.max.total.GB %>% tidyr::pivot_longer(!nsubj, names_to="software", values_to="max.total.GB.")

together.max.total.GB$software <- factor(together.max.total.GB$software, 
                                         levels = c("MRtrix","ModelArray") )


ggplot(together.max.total.GB, aes(fill=software, y=max.total.GB., x=nsubj)) + 
    geom_bar(position="dodge", stat="identity") + 
  scale_fill_manual(values= c("#909090", "#248B46") ) +
  geom_text(position = position_dodge(width = 50), 
            aes(y=max.total.GB.+50*0.05, label=sprintf("%0.2f", round(max.total.GB., digits=2)), hjust=0.4), 
            size=3.5, angle=45) + 
  ylim(0, 53) + 
  scale_x_continuous(breaks = c(summary.max$nsubj)) +
  theme_bw() + 
  theme(aspect.ratio = 1/2,
        text = element_text(size=15)) +
  xlab("Number of subjects") + 
  ylab("Max memory (GB)") + 
  ggtitle(paste0("Comparison: nthreads (ncores) = ",toString(nthreads)))



```

### sanity check of p values vs MRtrix's linear regression
#### number of shuffles = 100

```{r}

folder_mrtrix <- "/home/chenying/Desktop/fixel_project/data/data_from_josiane/for_fixelcfestats/stats_FDC/nsubj-938.nthreads-4.ftests.nshuffles-100.vmware.runMemProfiler.s-1sec.20210909-210227"
folder_fixelarray <- "/home/chenying/Desktop/fixel_project/data/data_from_josiane/results/ltn_FDC_n938_wResults_nfixel-0_20210908-135041"

# saved p values as .txt files via Confixel/notebooks/compare_MRtrix_ModelArray.py

f_fixelarray_raw_p <- paste0(folder_fixelarray, "/lm_model.p.value.txt")
f_fixelarray_p_fdr <- paste0(folder_fixelarray, "/lm_model.p.value.fdr.txt")
f_fixelarray_p_bonferroni <- paste0(folder_fixelarray, "/lm_model.p.value.bonferroni.txt")

fixelarray_raw_p <- read.csv(f_fixelarray_raw_p) %>% unlist() %>% as.numeric()
fixelarray_p_fdr <- read.csv(f_fixelarray_p_fdr) %>% unlist() %>% as.numeric()
fixelarray_p_bonferroni <- read.csv(f_fixelarray_p_bonferroni) %>% unlist() %>% as.numeric()


f_mrtrix_p_permu <- paste0(folder_mrtrix, "/uncorrected_pvalue_F1.txt")
f_mrtrix_p_permu_fwe <- paste0(folder_mrtrix, "/fwe_1mpvalue_F1.txt")

mrtrix_p_permu <- read.csv(f_mrtrix_p_permu) %>% unlist() %>% as.numeric()
mrtrix_p_permu <- 1 - mrtrix_p_permu
mrtrix_p_permu_fwe <- read.csv(f_mrtrix_p_permu_fwe) %>% unlist() %>% as.numeric()
mrtrix_p_permu_fwe <- 1 - mrtrix_p_permu_fwe


df <- data.frame(fixelarray_raw_p = fixelarray_raw_p,
                 fixelarray_p_fdr = fixelarray_p_fdr,
                 fixelarray_p_bonferroni = fixelarray_p_bonferroni,
                 mrtrix_p_permu = mrtrix_p_permu,
                 mrtrix_p_permu_fwe = mrtrix_p_permu_fwe)

cor.df <- cor(df)
cor.df
```
```{r}
f.all <- vector('list',5*5)
f.all <- matrix(f.all, nrow = 5)
#f.all <- matrix('list', nrow=5, ncol = 5)

# for (i in 1:5) {
#   for (j in 1:5) {
    
scatterPlotDF <- function(df,i,j) {
  cor.df <- cor(df)
  cor_str <- sprintf("%0.4f", round(cor.df[i,j], digits=4) ) 
    
  col.names <- colnames(df)
  
  myformula <- as.formula(paste0(col.names[j]," ~ ", col.names[i]))
  m <- lm(formula = myformula, data = df)
  
  print(m)
  # m %>% broom::tidy()
  m.glance <- m %>% broom::glance()   # model's statistics
  print("model p.value = ")
  print(m.glance$p.value, digits = 16)
  
  #f.all[[i,j]]
  f   <- ggplot(df, aes_string(x = col.names[i], 
                              y = col.names[j])) +
            #geom_point()  
            geom_bin2d(bins = 100) + 
            geom_smooth(method = "lm", se = FALSE) + 
            theme_bw() +
            theme(aspect.ratio = 1,
                  text = element_text(size=15)) + 
            xlab(colnames(df)[i]) + 
            ylab(colnames(df)[j]) + 
            ggtitle(paste0("corr = ",cor_str))
  
  
  f
}

f41 <- scatterPlotDF(df,4,1)
f42 <- scatterPlotDF(df,4,2)
f43 <- scatterPlotDF(df,4,3)

f51 <- scatterPlotDF(df,5,1)
f52 <- scatterPlotDF(df,5,2)
f53 <- scatterPlotDF(df,5,3)
    
  
p5 <- ggplot(df, aes(x = df[,5])) +
  geom_histogram(binwidth = 0.01) + 
  ylim(0,100)
h5 <- hist(df[,5], breaks = 101)
head(h5)   # has bins (0.01) that without any counts...


cor.df.longer <- cor.df %>% as.data.frame() %>% 
  rownames_to_column("whichpvalue") %>% 
  pivot_longer(-c("whichpvalue"), names_to = "sample", values_to = "corr")
f_corr <- ggplot(cor.df.longer, aes(x = sample, y = whichpvalue, fill = corr)) + 
  geom_raster() +
  scale_fill_distiller(palette = "YlOrRd")
f_corr
```




